---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hudr

<!-- badges: start -->
<!-- badges: end -->

This library contains an interface to
the US Department of Housing and Urban Development datasets. The goal
of this project is to provide an easy to use interface to access various open source
datasets provided by HUD. This library only contains a few datasets.


## Available Datasets

The APIs which this library interfaces with:

* Crosswalk
* Fair Markets Rent
* Income Limits
* Comprehensive Housing and Affordability Strategy

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("etam4260/hudr")
```

## Setup

The US Department of Housing and Urban Development requires users to gain an 
access key before querying their systems. You must go to https://www.huduser.gov/hudapi/public/register?comingfrom=1 and make an account.
From there you need to make a new token. Make sure to save the token somewhere as
you will only be able to view it once. You can now supply the 'key' argument.

To reduce the need to supply the key in every function call, you can use the 
setkey(key) function to save the key to the package environment.

```{r}
library(hudr)

# The function will not output anything. You can now run the other functions
# without supplying the key argument.
setkey("eyJ0eXAiOiJKV1QiLCJhbGciOiJSUzI1NiIsImp0aSI6IjM5OGJlNjBkNjYzMjM1ZmE2NzQxYWY5ZmViM2QzMDBkNDY3NTliYjgzMzhmNjJiZTE3ZDc4MmE0YWNhYjU2ZmMyMTIxMjM1MjJkYTVjNzY1In0.eyJhdWQiOiI2IiwianRpIjoiMzk4YmU2MGQ2NjMyMzVmYTY3NDFhZjlmZWIzZDMwMGQ0Njc1OWJiODMzOGY2MmJlMTdkNzgyYTRhY2FiNTZmYzIxMjEyMzUyMmRhNWM3NjUiLCJpYXQiOjE2NDI5ODg1MTgsIm5iZiI6MTY0Mjk4ODUxOCwiZXhwIjoxOTU4NTIxMzE3LCJzdWIiOiIyOTA3NCIsInNjb3BlcyI6W119.Ke0N8s797ohuGArbGb7rAMsLKDAWqP6mdItM8KjFQjHDMn8NYBazD8WopijiezC4wgV-n4n41NW4tSivV8yVow")
```


## Examples

### Crosswalk API
This is a basic example which shows you how to query the CROSSWALK dataset.

```{r}

# Type 7 corresponds to county level data. The query is a 5 digit fips code.
# The year and quarter specifies when these measurements were made. The key argument
# is the token you got from https://www.huduser.gov/hudapi/public/register?comingfrom=1
# This example does not have the key because we already set it up using setkey()
hudcw(type = 7, query = '22031', year = '2010', quarter = '4')
```

#### Crosswalk Data Fields

1) zip/fips/fipstract/CBSA/Congressional District/County Subdistrict > 
    * Zip, Tract, County, CD or CBSA code

2) year 
    * Value of year

3) quarter 
    * Quarter of the year

4) res_ratio  
    * The ratio of residential addresses in the ZIP – Tract, County, or CBSA part to the total number of residential addresses in the entire ZIP. (for type 1-5)
The ratio of residential addresses in the Zip, Tract, County, or CBSA - ZIP part to the total number of residential addresses in the entire Zip, Tract, County, or CBSA. (for type 6-10)

5) bus_ratio 
    * The ratio of business addresses in the ZIP – Tract, County, or CBSA part to the total number of business addresses in the entire ZIP. (for type 1-5)
The ratio of business addresses in the Tract, County, or CBSA – ZIP part to the total number of business addresses in the entire Tract, County, or CBSA. (for type 6-10)

6) oth_ratio 
    * The ratio of other addresses in the ZIP – Tract to the total number of other addresses in the entire ZIP. (for type 1-5)
The ratio of other addresses in the Tract, County, or CBSA – ZIP part to the total number of other addresses in the entire Tract, County, or CBSA. (for type 6-10)

7) tot_ratio 
    * The ratio of all addresses in the ZIP – Tract to the total number of all types of addresses in the entire ZIP. (for type 1-5)
The ratio of all addresses in the Tract, County, or CBSA-ZIP part to the total number of all types of addresses in the entire Tract, County, or CBSA. (for type 6-10)


### Fair Markets Rent API

This is a basic example to show you how to query the Fair Markets Rent API. Querying
a state will return a dataframe with rows representing counties within the state.
A county fips + subdivision 10 digit code will return a 1 row dataframe for that
county.

```{r}
# Querying the Fair Markets Rent dataset for Virginia in the year of 2021. This selected
# the first few rows and the first four columns.
head(hudfmr(query = 'VA', year = '2021'))[ ,c(1,2,3,4)]

# Querying the Fair Markets Rent dataset for a county in the year of 2017.
hudfmr(query = '0100199999', year = '2017')
```

#### Fair Markets Rent Data Fields
1) state/county or CBSA 
    * Name of the county if it is a county.

2) year 
    * Value of year

3) counties_msa 
    * Names of all counties belonging to the Metro Area if it is a Metro Area (MSA).

4) town_name 
    * Town name - applicable for North East regions

5) metro_status 
    * value will be "1" if it is a metropolitan county. Otherwise value will be "0".

6) metro_name 
    * Metro area name if metro_status is "1"

7) smallarea_status 
    * value will be "1" if it is a small area. Otherwise value will be "0".

8) Efficiency 
    * Efficiency FMR

9) One-Bedroom 
    * 1-bedroom FMR

10) Two-Bedroom 
    * 2-bedroom FMR

11) Three-Bedroom 
    * 3-bedroom FMR

12) Four-Bedroom 
    * 4-bedroom FMR

## Contributors

* Emmet Tam(https://github.com/etam4260)


## Disclaimers

* This interface uses the HUD User Data API but is not endorsed or certified by HUD User.
* The limit on the maximum number of API calls is 1200 queries a min.
* This is a WIP so please report any issues or bugs to: https://github.com/etam4260/hudr/issues
* License: MIT
* To get citation information for hudr in R, type citation(package = 'hudr')
* This is open source, so please fork and introduce some pull requests!
