---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# hudr

<!-- badges: start -->
<!-- badges: end -->

This library contains an interface to
the US Department of Housing and Urban Development datasets. 
The APIs which this library interfaces with:

CROSSWALKS

FAIR MARKETS RENT

INCOME LIMITS

COMPREHENSIVE HOUSING AND AFFORDABILITY STRATEGY

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("etam4260/hudr")
```

## Setup

The US Department of Housing and Urban Development requires users to gain an 
access key before querying their systems. You must go to https://www.huduser.gov/hudapi/public/register?comingfrom=1 and make an account.
From there you need to make a new token. Make sure to save the token somewhere as
you will only be able to view it once. You can now supply the 'key' argument
for each function.


## Examples

### Crosswalk API
This is a basic example which shows you how to query the CROSSWALK dataset.

```{r}
library(hudr)

# Type 7 corresponds to county level data. The query is a 5 digit fips code.
# The year and quarter specifies when these measurements were made. The key
# is the token you got from https://www.huduser.gov/hudapi/public/register?comingfrom=1
key = "eyJ0eXAiOiJKV1QiLCJhbGciOiJSUzI1NiIsImp0aSI6IjM5OGJlNjBkNjYzMjM1ZmE2NzQxYWY5ZmViM2QzMDBkNDY3NTliYjgzMzhmNjJiZTE3ZDc4MmE0YWNhYjU2ZmMyMTIxMjM1MjJkYTVjNzY1In0.eyJhdWQiOiI2IiwianRpIjoiMzk4YmU2MGQ2NjMyMzVmYTY3NDFhZjlmZWIzZDMwMGQ0Njc1OWJiODMzOGY2MmJlMTdkNzgyYTRhY2FiNTZmYzIxMjEyMzUyMmRhNWM3NjUiLCJpYXQiOjE2NDI5ODg1MTgsIm5iZiI6MTY0Mjk4ODUxOCwiZXhwIjoxOTU4NTIxMzE3LCJzdWIiOiIyOTA3NCIsInNjb3BlcyI6W119.Ke0N8s797ohuGArbGb7rAMsLKDAWqP6mdItM8KjFQjHDMn8NYBazD8WopijiezC4wgV-n4n41NW4tSivV8yVow"

hudcw(type = 7, query = '22031', year = '2010', quarter = '4', key = key)
```

#### Crosswalk Data Fields

zip/fips/fipstract/CBSA/Congressional District/County Subdistrict > 
Zip, Tract, County, CD or CBSA code

year >
Value of year

quarter >
Quarter of the year

res_ratio > 
The ratio of residential addresses in the ZIP – Tract, County, or CBSA part to the total number of residential addresses in the entire ZIP. (for type 1-5)
The ratio of residential addresses in the Zip, Tract, County, or CBSA - ZIP part to the total number of residential addresses in the entire Zip, Tract, County, or CBSA. (for type 6-10)

bus_ratio >
The ratio of business addresses in the ZIP – Tract, County, or CBSA part to the total number of business addresses in the entire ZIP. (for type 1-5)
The ratio of business addresses in the Tract, County, or CBSA – ZIP part to the total number of business addresses in the entire Tract, County, or CBSA. (for type 6-10)

oth_ratio >
The ratio of other addresses in the ZIP – Tract to the total number of other addresses in the entire ZIP. (for type 1-5)
The ratio of other addresses in the Tract, County, or CBSA – ZIP part to the total number of other addresses in the entire Tract, County, or CBSA. (for type 6-10)

tot_ratio >
The ratio of all addresses in the ZIP – Tract to the total number of all types of addresses in the entire ZIP. (for type 1-5)
The ratio of all addresses in the Tract, County, or CBSA-ZIP part to the total number of all types of addresses in the entire Tract, County, or CBSA. (for type 6-10)


### Fair Markets Rent API

This is a basic example to show you how to query the Fair Markets Rent API. Query
a state will return a dataframe with rows representing counties within the state.
A county fips + subdivision 10 digit code will return a 1 row dataframe for that
county.

```{r}
key = "eyJ0eXAiOiJKV1QiLCJhbGciOiJSUzI1NiIsImp0aSI6IjM5OGJlNjBkNjYzMjM1ZmE2NzQxYWY5ZmViM2QzMDBkNDY3NTliYjgzMzhmNjJiZTE3ZDc4MmE0YWNhYjU2ZmMyMTIxMjM1MjJkYTVjNzY1In0.eyJhdWQiOiI2IiwianRpIjoiMzk4YmU2MGQ2NjMyMzVmYTY3NDFhZjlmZWIzZDMwMGQ0Njc1OWJiODMzOGY2MmJlMTdkNzgyYTRhY2FiNTZmYzIxMjEyMzUyMmRhNWM3NjUiLCJpYXQiOjE2NDI5ODg1MTgsIm5iZiI6MTY0Mjk4ODUxOCwiZXhwIjoxOTU4NTIxMzE3LCJzdWIiOiIyOTA3NCIsInNjb3BlcyI6W119.Ke0N8s797ohuGArbGb7rAMsLKDAWqP6mdItM8KjFQjHDMn8NYBazD8WopijiezC4wgV-n4n41NW4tSivV8yVow"

hudfmr(query = 'VA', year = '2021', key = key)

hudfmr(query = '0100199999', year = '2017', key = key)
```

#### Fair Markets Rent Data Fields
state/county or CBSA >
Name of the county if it is a county.

year >
Value of year

counties_msa >
Names of all counties belonging to the Metro Area if it is a Metro Area (MSA).

town_name >
Town name - applicable for North East regions

metro_status >
value will be "1" if it is a metropolitan county. Otherwise value will be "0".

metro_name >
Metro area name if metro_status is "1"

smallarea_status >
value will be "1" if it is a small area. Otherwise value will be "0".

Efficiency >
Efficiency FMR

One-Bedroom >
1-bedroom FMR

Two-Bedroom >
2-bedroom FMR

Three-Bedroom >
3-bedroom FMR

Four-Bedroom >
4-bedroom FMR

