---
title: "Community-Development-Block-Grant"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Community-Development-Block-Grant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Program Overview

"The Community Development Block Grant (CDBG) Program has Disaster Recovery
grants to rebuild the affected areas and provide crucial seed money to start the
recovery process. These flexible grants help cities, counties, and States
recover from Presidentially declared disasters, especially in low-income areas,
subject to availability of supplemental appropriations. Since CDBG Disaster
Recovery (CDBG-DR) assistance may fund a broad range of recovery activities, HUD
can help communities and neighborhoods that otherwise might not recover due to
limited resources (https://www.hud.gov/program_offices/comm_planning/cdbg-dr)."

The datasets which this function deals with are located at
https://drgr.hud.gov/public/data_downloads.html?programName=DR%20CDBG. 
The first two datasets are the same, but are organized in a different manner.

Querying the HUD Exchange website for data is a bit different than HUD User.
These datasets do not need a key.

Files:
  1) DR CDBG CDBG-DR Financial Report by Appropriation
  2) DR CDBG CDBG-DR Financial Report by Grantee
  3) DR CDBG CDBG-DR Financial Report Monthly Summary
  4) DR CDBG CDBG-DR Performance by Activity

### Example

#### CDBG-DR Financial Report by Appropriation

This is an example grabbing the first file, DR CDBG CDBG-DR Financial Report
by Appropriation, and displaying the first 6 rows.
```{r, eval=FALSE}
library(hudr)
head(hud_cdbg(file = 1))
```

#### Output Data

1) Appropriation Abbreviation

2) Appropriation 

3) Grantee	

4) Grant	

5) Grant Status	

6) Grant $ (Grant Level)	

7) Grant Obligated $	

8) Obligated %	

9) Grant Disbursed $	

10) Disbursed %	

11) Remaining Balance

#### CDBG CDBG-DR Financial Report by Grantee

This is an example grabbing the second file, DR CDBG CDBG-DR Financial Report by
Grantee, and displaying the first 6 rows.
```{r, eval=FALSE}
head(hud_cdbg(file = 2))
```

#### Output Data

1) Grantee State

2) Appropriation		

3) Grantee	Grant	

4) Grant Status	

5) Grant $ (Grant Level)	

6) Grant Obligated $	

7) Obligated %	

8) Grant Disbursed $	

9) Disbursed %	

10) Remaining Balance

#### CDBG-DR Financial Report Monthly Summary

This is an example grabbing the third file, DR CDBG CDBG-DR Financial Report
Monthly Summary, and displaying the first 6 rows. The data provides grants that
were given in the last month.
```{r, eval=FALSE}
head(hud_cdbg(file = 3))
```

#### Output Data

1) Grantee State

2) Grantee	

3) Grant	

4) Grant $ (deprecated)

5) Activity Budget $	

6) % Budget	

7) Activity Obligation $	

8) % Obligated	

9) Activity Drawn $	

10) Activity Budget Disbursed %	

11) Remaining Balance

#### DR CDBG CDBG-DR Performance by Activity

This is an example grabbing the fourth file, DR CDBG CDBG-DR Performance by
Activity, and displaying the first 6 rows.
```{r, eval=FALSE}
head(hud_cdbg(file = 4))
```

#### Output Data

1) Appropriation		

2) Grantee	

3) Grant	

4) Activity Responsible Org	

5) Activity Type	

6) Activity Number	

7) Activity Title	

8) Activity Status	

9) Activity Measure Category	

10) Low Projected

11) Low Actual

12) Low Actual to Projected %

13) Mod Projected

14) Mod Actual

15) Mod Actual to Projected %

16) Total Projected

17) Total Actual

18) Total Actual to Projected %
