---
title: "Community-Development-Block-Grant"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Community-Development-Block-Grant}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### About

"The Community Development Block Grant (CDBG) Program has Disaster Recovery grants to rebuild the affected areas and provide crucial seed money to start the recovery process. These flexible grants help cities, counties, and States recover from Presidentially declared disasters, especially in low-income areas, subject to availability of supplemental appropriations. Since CDBG Disaster Recovery (CDBG-DR) assistance may fund a broad range of recovery activities, HUD can help communities and neighborhoods that otherwise might not recover due to limited resources (https://www.hud.gov/program_offices/comm_planning/cdbg-dr)."

The datasets which this function deals with are located at https://drgr.hud.gov/public/data_downloads.html?programName=DR%20CDBG
The first two datasets are the same, but are organized in a different manner.

Querying the HUD Exchange website for data is a bit different than HUD User. These functions
directly download the data into memory. Furthermore, these datasets do not need a key.

### Example

#### CDBG-DR Financial Report by Appropriation

This is an example grabbing the first file, DR CDBG CDBG-DR Financial Report
by Appropriation, and displaying the first 6 rows.
```{r}
library(hudr)
head(hudcdbg(file = 1))
```

#### Data Fields

1) Appropriation Abbreviation

2) Appropriation 

3) Grantee	

4) Grant	

5) Grant Status	

6) Grant $ (Grant Level)	

7) Grant Obligated $	

8) Obligated %	

9) Grant Disbursed $	

10) Disbursed %	

11) Remaining Balance

#### CDBG CDBG-DR Financial Report by Grantee

This is an example grabbing the second file, DR CDBG CDBG-DR Financial Report by Grantee, 
and displaying the first 6 rows.
```{r}
head(hudcdbg(file = 2))
```

#### Data Fields

Grantee State

Appropriation		

Grantee	Grant	

Grant Status	

Grant $ (Grant Level)	

Grant Obligated $	

Obligated %	

Grant Disbursed $	

Disbursed %	

Remaining Balance


#### CDBG-DR Financial Report Monthly Summary

This is an example grabbing the third file, DR CDBG CDBG-DR Financial Report Monthly Summary, 
and displaying the first 6 rows. The data provides grants that were given in the last month.
```{r}
head(hudcdbg(file = 3))
```

#### Data Fields

Grantee State

Grantee	

Grant	

Grant $ (deprecated)

Activity Budget $	

% Budget	

Activity Obligation $	

% Obligated	

Activity Drawn $	

Activity Budget Disbursed %	

Remaining Balance

#### DR CDBG CDBG-DR Performance by Activity

This is an example grabbing the fourth file, DR CDBG CDBG-DR Performance by Activity, 
and displaying the first 6 rows.
```{r}
head(hudcdbg(file = 4))
```

#### Data Fields

Appropriation		

Grantee	

Grant	

Activity Responsible Org	

Activity Type	

Activity Number	

Activity Title	

Activity Status	

Activity Measure Category	

Low			
Projected   Actual	Actual to Projected %

Mod
Projected   Actual	Actual to Projected %

Total
Projected   Actual	Actual to Projected %
