---
title: "Comprehensive-Housing-and-Affordability-Strategy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comprehensive-Housing-and-Affordability-Strategy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Comprehensive Housing and Affordability Strategy API

This is a basic example showing you how to query the COMPREHENSIVE HOUSING AND AFFORDABILITY STRATEGY API.

This function returns a single row dataframe containing information about household
income levels, occupancy, housing problems, and burden for a 'region' 

```{r}
library(hudr)

# Setting the key.
hudsetkey("eyJ0eXAiOiJKV1QiLCJhbGciOiJSUzI1NiIsImp0aSI6IjM5OGJlNjBkNjYzMjM1ZmE2NzQxYWY5ZmViM2QzMDBkNDY3NTliYjgzMzhmNjJiZTE3ZDc4MmE0YWNhYjU2ZmMyMTIxMjM1MjJkYTVjNzY1In0.eyJhdWQiOiI2IiwianRpIjoiMzk4YmU2MGQ2NjMyMzVmYTY3NDFhZjlmZWIzZDMwMGQ0Njc1OWJiODMzOGY2MmJlMTdkNzgyYTRhY2FiNTZmYzIxMjEyMzUyMmRhNWM3NjUiLCJpYXQiOjE2NDI5ODg1MTgsIm5iZiI6MTY0Mjk4ODUxOCwiZXhwIjoxOTU4NTIxMzE3LCJzdWIiOiIyOTA3NCIsInNjb3BlcyI6W119.Ke0N8s797ohuGArbGb7rAMsLKDAWqP6mdItM8KjFQjHDMn8NYBazD8WopijiezC4wgV-n4n41NW4tSivV8yVow")

# Querying the Comprehensive Housing and Affordability Strategy dataset for a state in the years of 2014-2018.
# Return the first 6 rows. There are specific date ranges that are allowed. Please check the specifications.
# https://www.huduser.gov/portal/dataset/chas-api.html

# This displays the first 9 columns
hudchas(type = '2', stateId ='56', year = '2014-2018')[,c(1,2,3,4,5,6,7,8,9)]
```

#### Comprehensive Housing and Affordability Strategy Data Fields

I recommend checking out https://www.huduser.gov/portal/dataset/chas-api.html for more
details on this as there are too many fields to display.

