---
title: "Covid-19-and-Evictions-Crosswalk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Covid-19-and-Evictions-Crosswalk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<i>Sample provided by <b>Shuyu Jin (geojsy@umd.edu)</b></i>


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

### Overview
In the United States, housing advocates and policymakers are concerned that a sudden disruption in household finances caused by the economic crisis related to the COVID-19 pandemic will lead to a wave of evictions by tenants unable to pay their rents [(*Cowin, Martin, and Stevens, 2020*)](https://www.clevelandfed.org/en/newsroom-and-events/publications/community-development-briefs/db-20200717-measuring-evictions-during-the-covid-19-crisis.aspx). 

Viewing [*2016 eviction rankings*](https://evictionlab.org/rankings/#/evictions?r=United%20States&a=0&d=evictionRate&lang=en) for locations across the U.S., Richmond has maintained a high eviction rate over the past few years. To understand how the pandemic is affecting evictions, I collected weekly eviction filing data for Richmond, which are provided by [*Princeton’s Eviction Lab*](https://evictionlab.org/eviction-tracking/). 

The mapping below uses ggplot and the crosswalk files to display COVID-19 cases as well as evictions based on zip codes and census tracts.


### Loading in the necessary libraries
```{r message=FALSE}
library(rhud)
library(sf)
library(dplyr)
library(scales)
library(ggplot2)
options (digits=20)
```
To improve performance, use tigris caching.

```{r message=FALSE}
options(tigris_use_cache = TRUE)
```


### Reading in data

```{r message=FALSE}
library(readr)
richmond_monthly_2020_2021 <- read_csv("Data/richmond_monthly_2020_2021.csv")
VA_COVID_19_PublicUseDataset_ZIPCode <- read_csv("Data/VA-COVID-19-PublicUseDataset-ZIPCode.csv")

# Using tigris package to download Richmond shapefiles
richmond_shapefile_zctas = tigris::zctas(cb = FALSE, starts_with = "23")
richmond_shapefile_tracts = tigris::tracts(state = "51", county = "760", cb = TRUE)
```

Cleaning date objects
```{r message=FALSE}
library(lubridate)
VA_COVID_19_PublicUseDataset_ZIPCode$date <- lubridate::mdy(VA_COVID_19_PublicUseDataset_ZIPCode$`Report Date`)

VA_COVID_19_PublicUseDataset_ZIPCode$monthYear <- paste(as.character( lubridate::month(VA_COVID_19_PublicUseDataset_ZIPCode$date)), "-", as.character( lubridate::year(VA_COVID_19_PublicUseDataset_ZIPCode$date)), sep="" )

VA_COVID_19_PublicUseDataset_ZIPCode$monthYear <- my(VA_COVID_19_PublicUseDataset_ZIPCode$monthYear)

VA_COVID_19_PublicUseDataset_ZIPCode$`Number of Cases` <- as.integer(VA_COVID_19_PublicUseDataset_ZIPCode$`Number of Cases`)

# Aggregating a new data frame with number of cases, monthYear, and ZIP code
covidCases <- aggregate(`Number of Cases` ~ monthYear + `ZIP Code`, data = VA_COVID_19_PublicUseDataset_ZIPCode, FUN = sum)
```

Aggregating racial majority based on eviction fillings data
```{r message=FALSE}
eviction_race <- aggregate(`filings_2020` ~ month + GEOID + racial_majority, data = richmond_monthly_2020_2021, FUN = sum)
eviction_no_race <- aggregate(`filings_2020` ~ month + GEOID, data = richmond_monthly_2020_2021, FUN = sum)
```

Aggregating and classifying COVID cases and eviction data based on spatial and temporal condition
```{r message=FALSE}
covidCases_spatial <- aggregate(`Number of Cases` ~ `ZIP Code`, FUN = sum, data = covidCases)
covidCases_temporal <- aggregate(`Number of Cases` ~ monthYear, FUN = sum, data = covidCases)

eviction_no_race_spatial <- aggregate(filings_2020 ~ GEOID, FUN = sum, data = eviction_no_race)
eviction_no_race_temporal <- aggregate(filings_2020 ~ month, FUN = sum, data = eviction_no_race)
eviction_no_race_temporal$month <- my(eviction_no_race_temporal$month)
```


### Mapping COVID cases and Eviction Filings by ZIP Codes

```{r message=FALSE}
richmond <- inner_join(richmond_shapefile_zctas, eviction_no_race_spatial, by = c("ZCTA5CE20" = "GEOID" ))
richmond <- inner_join(richmond, covidCases_spatial, by = c("ZCTA5CE20" = "ZIP Code" ))
```

```{r message=FALSE}
ggplot() +
  geom_sf(data = richmond, aes(fill= `Number of Cases`), size = 0.1, colour = "grey") + 
  scale_fill_gradient2(low = "grey", mid = "white", high = "darkred", labels = comma) +
  ggtitle("Covid-19 Case in Richmond, VA during 2020-2021, by zipcode") +
  coord_sf() +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white"))
```

```{r message=FALSE}
ggplot() +
  geom_sf(data = richmond, aes(fill= `filings_2020`), size = 0.1, colour = "grey") + 
  scale_fill_gradient2() +
  ggtitle("Eviction Filings in Richmond, VA during 2020-2021, by zipcode") +
  coord_sf() +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white"))
```

```{r}
richmondTemporal <- inner_join(eviction_no_race_temporal, covidCases_temporal, by = c("month" = "monthYear"))
colnames(richmondTemporal) <- c("Date", "Eviction","CovidCase")
```


```{r}
options(scipen = 999)
ggplot(richmondTemporal, aes(Date)) + 
  geom_line(aes(y = CovidCase),color = "red") +
theme(axis.line = element_line(colour = "black"),
      panel.background = element_blank())
```

```{r}
ggplot(richmondTemporal, aes(Date)) + 
  geom_line(aes(y = Eviction), col = "blue") +
  theme(axis.line = element_line(colour = "black"),
        panel.background = element_blank())
```

```{}
Although the number of COVID cases and eviction filings does not appear to be positively related, in particular, eviction filings have almost returned to their pre-pandemic levels but increased in the early month of 2021 in Richmond. It provides an idea of what factors lead to changes in eviction.
```



### Mapping COVID cases and Eviction Filings by Census Tracts

```{r message=FALSE}
richmond$ZCTA5CE20 <- as.numeric(richmond$ZCTA5CE20)
```

Crosswalk zipcode to census tract
```{r message=FALSE}
# Converting from a string to a dataframe to calculate
a = as.data.frame(richmond)
covidcases_cw <- crosswalk(data = a, geoid = "zip", geoid_col = "ZCTA5CE20",
          cw_geoid = "tract", cw_geoid_col = "Number of Cases", method = "tot")

evi_cw <- crosswalk(data = a, geoid = "zip", geoid_col = "ZCTA5CE20",
          cw_geoid = "tract", cw_geoid_col = "filings_2020", method = "tot")
```
Aggregating date with census tract level
```{r message=FALSE}
evi_tract <- aggregate(filings_2020 ~ tract, FUN = sum, data = evi_cw)
evi_tract$filings_2020 <- as.integer(evi_tract$filings_2020)

cov_tract <- aggregate(`Number of Cases` ~ tract, FUN = sum, data = covidcases_cw)
cov_tract$`Number of Cases` <- as.integer(cov_tract$`Number of Cases`)
```


```{r message=FALSE}
evi <- inner_join(richmond_shapefile_tracts, evi_tract, by = c("GEOID" = "tract"))
covid <- inner_join(richmond_shapefile_tracts, cov_tract, by = c("GEOID" = "tract"))
```

```{r message=FALSE}
ggplot() +
  geom_sf(data = covid, aes(fill= `Number of Cases`), size = 0.1, colour = "grey") + 
  scale_fill_gradient2(low = "grey", high = "darkred", labels = comma) +
  ggtitle("Covid Case in Richmond, VA during 2020-2021, by census tract") +
  coord_sf() +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white"))
```


```{r message=FALSE}
ggplot() +
  geom_sf(data = evi, aes(fill= `filings_2020`), size = 0.1, colour = "grey") + 
  scale_fill_gradient2() +
  ggtitle("Eviction Filings in Richmond, VA during 2020-2021, by census tract") +
  coord_sf() +
  theme(plot.title = element_text(face = "bold"),
        axis.text.x = element_blank(), 
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
    panel.background = element_rect(fill = "white"))
```


## Observations
A noteworthy point, the number of COVID and eviction filings is not significantly correlated. In fact, by 2021, Richmond remained 25% below the historical average [*(Hepburn, Jin, & Fish et al., 2022)*](https://evictionlab.org/us-eviction-filing-patterns-2021/). To answer why the number of eviction filings in Richmond has been changed during the pandemic, combing with time, the CDC moratorium expiration might be a significant reason. Since during September 4, 2020, and July 31, 2021, [*CDC’s moratorium*](https://www.cdc.gov/media/releases/2021/s0803-cdc-eviction-order.html) on evictions for rent assistance for tenants and landlords helped prevent at least 1.55 million eviction filings nationwide [*(Rangel, Haas, and Lemmerman et al., 2021)*](https://evictionlab.org/eleven-months-cdc/).

Meanwhile, crossing data from ZIP Codes to census tracts yielded less accurate results. We got some discrete or even N/A results with COVID cases by census tracts. Therefore, it should be noted that there are considerations when crosswalking data to smaller geographies, particularly census tracts. The fundamental reason should be the ZIP Codes cross-cut numerous tracts that do not follow the boundaries closely [*(Din and Wilson, 2020)*]( https://www.huduser.gov/portal/periodicals/cityscpe/vol22num1/ch12.pdf). 



## Reference
CDC. "CDC Issues Eviction Moratorium Order in Areas of Substantial and High Transmission." https://www.cdc.gov/media/releases/2021/s0803-cdc-eviction-order.html (Accessed 05/12/2022)

Din, Alexander, and Ron Wilson. "Crosswalking ZIP Codes to Census Geographies." Cityscape 22, no. 1 (2020): 293-314. (Accessed 06/01/2022)

Jasmine Rangel, Jacob Haas, Emily Lemmerman, et al. “Preliminary Analysis: 11 Months of the CDC Moratorium.” Eviction Lab. https://evictionlab.org/eleven-months-cdc/. (Accessed 05/12/2022)

Peter Hepburn, Olivia Jin, Joe Fish, et al. “Preliminary Analysis: Eviction Filing Patterns in 2021.” Eviction Lab. https://evictionlab.org/us-eviction-filing-patterns-2021/. (Accessed 05/14/2022)
