---
title: "Crosswalk-a-Dataset"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Crosswalk-a-Dataset}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Obtaining a sample dataset

For those who need a more powerful method for crosswalking semi-large datasets 
and allocating fields based on 
the provided ratios, the crosswalk() function is available.

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE}
library(rhud)
options (digits=4)

sample <- data.frame(pop = c(151049, 103609, 52276),
                     county = c("24043", "24045", "24047"))

head(sample)
```

This sample dataset contains population information
for the counties of Washington, Wicomico, and Worchester in Maryland in the
year 2019 according to the US Census Bureau.

<br><br><br>

## Merging with a crosswalk file

Lets say we wanted to know the population at a zipcode level (there is likely 
already a dataset for this), but utilizing the crosswalk files in this case can
provide a little insight into its effectiveness as truth answers are available
to compare. 

In this case we care about using the county to zip code crosswalk measurements 
from the year 2019 in the first quarter of the year.

To merge the dataset with its associated crosswalk measurement use:

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE}

crosswalk(data = sample, geoid = "county", geoid_col = "county", 
          cw_geoid = "zip", cw_geoid_col = NA, method = NA, year = 2019
          , quarter = 1)

```

As seen in the data above, each zipcode associated with a county is assigned the
same population value.

<br><br><br>

## Merging and applying a crosswalk method

To utilize an allocation method provided by the crosswalk files, specify
the method and cw_geoid_col arguments. In this case we want to allocate
the county population levels to a zip code level using the method
based on the ratio of residential addresses. 

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE}

crosswalk(data = sample, geoid = "county", geoid_col = "county", 
          cw_geoid = "zip", cw_geoid_col = "pop", method = "res", year = 2019
          , quarter = 1)
```

As seen in the population field, it is now partitioned out based on
the ratio of residential addresses for each zip code that comprises a county.
