---
title: "Fair-Markets-Rent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fair-Markets-Rent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Program Overview

According to ("What Is Fair Market Rent? | RentData.org"), Fair Market Rent
refers to the monthly rent, including utilities (excluding phone), for privately
owned, existing, adequate, safe, and sanitary rental accommodation of a moderate
(non-luxury) kind with appropriate facilities.

The Department of Housing and Urban Development (HUD) established the Fair
Market Rent rate to be used as a foundation for funding federal housing
assistance programs like the Housing Choice Voucher Program.

Each county and metro area in the United States has its own set of Fair Market
Rent pricing. The prices range from zero-bedroom studios or efficiency units to
four-bedroom residences.

The 40th percentile of rentals paid by recent movers (renters who relocated
within the last 24 months) in a specific FMR location is defined as Fair Market
Rent (FMR). FMR rent costs are slightly lower than the median.


According to (Fair Market Rents | HUD USER), Fair Market Rents (FMRs) are used
to calculate payment standard amounts for the Housing Choice Voucher program,
initial renewal rents for some expiring project-based Section 8 contracts,
initial rents for housing assistance payment (HAP) contracts in the Moderate
Rehabilitation Single Room Occupancy program (Mod Rehab), rent ceilings for
rental units in both the HOME Investment Partnerships program and the Emergency
Solution Grants program, and rent ceilings for rental units in both the HOME
Investment Partnerships program and the Emergency Solution Grant.

Furthermore, the United States Department of Housing and Urban Development (HUD)
calculates FMRs for metropolitan areas defined by the Office of Management and
Budget (OMB), some HUD-designated subdivisions of OMB metropolitan areas, and
each non-metropolitan county on an annual basis. FMRs must be posted at least 30
days before they go into effect, and they must go into force at the beginning of
the federal fiscal year (usually October 1). Fair Market Rentals are estimates
of the 40th percentile gross rents for standard quality units within a
metropolitan region or non-metropolitan county, as described in 24 CFR 888.113.

### Examples

These are basic examples to show you how to query the Fair Markets Rent API. Before
looking at the outputted data I RECOMMEND first looking at [Input Arguments][]
as well as [Output Data][] located at the bottom of the page.

**DISCLAIMER:** The output tables are too large to be displayed here so we choose to
leave them out.

Querying a county or cbsa code that are considered small areas will 
return data at a zip code level. If they are not small areas, then it will 
give the measurement for that entire county or cbsa. 

Lets say we are working with county level data.

When mixing them together (counties that are small areas and not small areas)
those counties that are not small areas will have an NA in the zip_code, 
meaning this measurement is defined at a county level. Those counties that 
are small areas will return data in the zip_code field.

#### US States
This is an example to show you how to query for all states in the US.

```{r, eval=FALSE}
library(hudr)
hud_nation_states_territories()
```

#### US Counties 
This is an example to show you how to query for all counties in MD,
Virginia, and California, respectively.

```{r, eval=FALSE}
hud_state_counties("MD")

hud_state_counties("Virginia")

hud_state_counties("6")
```

#### US Metropolitan Areas
This is an example to show you how to query for metropolitan areas in Wyoming
and New York.
```{r, eval=FALSE}
hud_state_metropolitan(c("WY", "NY"))
```

#### State Level Fair Markets Rent
This is an example which shows you how to query the Fair Markets Rent API for a
state. You can use the state abbreviation, state name, or state fips code. It
will return all counties in the state and their Fair Markets Rents as well
as all metroareas in the state and their Fair Markets Rent.
```{r, eval=FALSE}
# These functions gets FMR data for the state of Virginia in 2021.
hud_fmr(query = 'VA', year = '2021')

hud_fmr(query= "Virginia", year = '2021')

hud_fmr(query= "51", year = '2021')
```

#### County Level Fair Markets Rent
This is an example which shows you how to query the Fair Markets Rent on a
county level basis. These are defined with a 2 digit state code + 3 digit county
code. You also need a 99999 code added onto the end. We currently don't know why
the extra 99999 is included and have not found any cases where its different. I
recommend taking a look at [US Counties][] to determine what counties are
available.
```{r, eval=FALSE}
# Getting a county requires a 2 digit state fipscode + 3 digit county fipscode + 99999
hud_fmr(query = '0100199999', year = '2017')
```

#### Small Areas Fair Markets Rent
This is an example which shows you how to query the Small Areas Fair Markets Rent.
Small Areas are defined as metropolitan and micropolitan areas which are defined in
CBSA codes. Currently we know how to query for metropolitan, but not sure for 
micropolitan yet. I recommend
taking a look at [US Metropolitan Areas][] to determine what metroareas
are available.
```{r, eval=FALSE}
# Gets FMR data for METRO area.
hud_fmr(query = "METRO47900M47900", year=c(2018))
```

### Input Arguments

**QUERY**: Can provide either a 10 digit FIPS code which is almost always
  state fips + county fips + 99999, or state abbreviation. 
  Can also provide a CBSA code. You are only allowed to query for metropolitan areas.
  
  Run hud_states() to get a list of counties.
  
  Run hud_metropolitan("MD") to get a list of metropolitan areas in MD.
  
  Run hud_counties("MD") to get list of counties in MD.
  
  * query = 'METRO12700M12700'
  * query = 'MD'
  * query = '5100199999'

**YEAR**: Year range of the data to retrieve. Defaults to current year. 

  * year = c(2019, 2018, 2021)
  * year = c(2016)
  * year = 2021

**KEY**: The API key.

  * key = "wqokqo2138jdi13wfwwfwcytjyr"
  
### Output Data 

1) **state/county or CBSA**
    * Name of the county if it is a county.


2) **year** 
    * Value of year


3) **counties_msa**
    * Names of all counties belonging to the Metro Area if it is a Metro Area
    (MSA).


4) **town_name**
    * Town name - applicable for North East regions


5) **metro_status**
    * value will be "1" if it is a metropolitan county. Otherwise value will be
    "0".


6) **metro_name** 
    * Metro area name if metro_status is "1"


7) **smallarea_status** 
    * value will be "1" if it is a small area. Otherwise value will be "0".


8) **Efficiency(A studio apartment)** 
    * Efficiency FMR in US Dollars


9) **One-Bedroom** 
    * 1-bedroom FMR in US Dollars


10) **Two-Bedroom** 
    * 2-bedroom FMR in US Dollars


11) **Three-Bedroom** 
    * 3-bedroom FMR in US Dollars


12) **Four-Bedroom** 
    * 4-bedroom FMR in US Dollars
    

### References

"What Is Fair Market Rent? | RentData.org." Www.rentdata.org,
<br />&nbsp;&nbsp;&nbsp;&nbsp; www.rentdata.org/articles/what-is-fair-market-rent. Accessed 18 Feb. 2022.

"Fair Market Rents | HUD USER." Huduser.gov, U.S Department of Housing and Urban
<br />&nbsp;&nbsp;&nbsp;&nbsp; Development, 2017, www.huduser.gov/portal/datasets/fmr.html.
