---
title: "Fair-Markets-Rent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fair-Markets-Rent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### About

"Fair Market Rent is the rent amount, including utilities (except telephone), to rent privately owned, existing, decent, safe and sanitary rental housing of modest (non-luxury) nature with suitable amenities.

The Fair Market Rent rate established by Department of Housing and Urban Development (HUD) to be used as a basis for paying federal housing assistance programs like the Housing Choice Voucher Program.

Fair Market Rent prices are set for each county and metro area across the country. Rates are set for zero-bedroom studios or efficiencies all the way up to 4-bedroom homes.

Fair Market Rent (FMR) is defined as the 40th percentile of rents paid by recent movers (renters who moved in the last 24 months) in a given FMR area. This means FMR rent prices are slightly below the median price (https://www.rentdata.org/articles/what-is-fair-market-rent)."

query: Can provide either a 10 digit FIPS code including county subdivision, or state abbreviation.

year: Gets the year that this data was recorded. Can specify multiple years. Default is the latest year.

### Example

This is a basic example to show you how to query the Fair Markets Rent API. 

Querying a state will return a dataframe with rows representing counties within the state.
A (state fip + county fip + county subdivision) 10 digit code will return a 1 row dataframe for that
county.

```{r, eval=FALSE}
library(hudr)

# This selects the first few rows and the first four columns.
hud_fmr(query = 'VA', year = '2021')

# Getting a county requires a county fip + state fip + census tract.
hud_fmr(query = '0100199999', year = '2017')
```

### Data Fields

1) state/county or CBSA 
    * Name of the county if it is a county.


2) year 
    * Value of year


3) counties_msa 
    * Names of all counties belonging to the Metro Area if it is a Metro Area (MSA).


4) town_name 
    * Town name - applicable for North East regions


5) metro_status 
    * value will be "1" if it is a metropolitan county. Otherwise value will be "0".


6) metro_name 
    * Metro area name if metro_status is "1"


7) smallarea_status 
    * value will be "1" if it is a small area. Otherwise value will be "0".


8) Efficiency(A studio apartment) 
    * Efficiency FMR in US Dollars


9) One-Bedroom 
    * 1-bedroom FMR in US Dollars


10) Two-Bedroom 
    * 2-bedroom FMR in US Dollars


11) Three-Bedroom 
    * 3-bedroom FMR in US Dollars


12) Four-Bedroom 
    * 4-bedroom FMR in US Dollars
