---
title: "Fair-Markets-Rent"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fair-Markets-Rent}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Program Overview

"Fair Market Rent is the rent amount, including utilities (except telephone),
to rent privately owned, existing, decent, safe and sanitary rental housing of 
modest (non-luxury) nature with suitable amenities.

The Fair Market Rent rate established by Department of Housing and Urban 
Development (HUD) to be used as a basis for paying federal housing assistance
programs like the Housing Choice Voucher Program.

Fair Market Rent prices are set for each county and metro area across the
country. Rates are set for zero-bedroom studios or efficiency all the way
up to 4-bedroom homes.

Fair Market Rent (FMR) is defined as the 40th percentile of rents paid by 
recent movers (renters who moved in the last 24 months) in a given FMR area. 
This means FMR rent prices are slightly below the median price 
("What Is Fair Market Rent? | RentData.org")."


"Fair Market Rents (FMRs) are used to determine payment standard amounts for the
Housing Choice Voucher program, to determine initial renewal rents for some
expiring project-based Section 8 contracts, to determine initial rents for
housing assistance payment (HAP) contracts in the Moderate Rehabilitation Single
Room Occupancy program (Mod Rehab), rent ceilings for rental units in both the
HOME Investment Partnerships program and the Emergency Solution Grants program,
calculation of maximum award amounts for Continuum of Care recipients and the
maximum amount of rent a recipient may pay for property leased with Continuum of
Care funds, and calculation of flat rents in Public Housing units. The U.S.
Department of Housing and Urban Development (HUD) annually estimates FMRs for
Office of Management and Budget (OMB) defined metropolitan areas, some HUD
defined subdivisions of OMB metropolitan areas and each nonmetropolitan county.
42 USC 1437f requires FMRs be posted at least 30 days before they are effective
and that they are effective at the start of the federal fiscal year (generally
October 1). Fair Market Rents, as defined in 24 CFR 888.113 are estimates of
40th percentile gross rents for standard quality units within a metropolitan
area or nonmetropolitan county ("Fair Market Rents | HUD USER")."

### Example

This is a basic example to show you how to query the Fair Markets Rent API. 

Querying a state will return a dataframe with rows representing counties 
within the state. A (state fip + county fip + county subdivision) 10 digit code
will return a 1 row dataframe for that county.

For records corresponding to parts of or entire place entities that do not
overlap any 2010 urban area, the urban area code is 99999, the urban area name
is “Not in a 2010 urban area”, and the urban area population, housing unit
count, total area, and land area values are null. The percent values relating to
the urban area are also null.

```{r, eval=FALSE}
library(hudr)

# This gets FMR data for the state of Virginia in 2021.
hud_fmr(query = 'VA', year = '2021')

# Getting a county requires a county fip + state fip + area code.
# The 99999 corresponding to parts of or entire place entities that do not
# overlap any 2010 urban area.
hud_fmr(query = '0100199999', year = '2017')

# I recommend first running to first determine full county codes that can
# be queried for a state. 
hud_counties('VA') 

# You can also check for CBSA codes which can also be supplied in hud_fmr()
# query field.
hud_metropolitan('VA') 
```

### Input Arguments

QUERY: Can provide either a 10 digit FIPS code including county 
subdivision, or state abbreviation. Can also provide a CBSA code.

YEAR: Gets the year that this data was recorded. Can specify multiple years.
Defaults to the previous year.

KEY: The API key.

### Output Data 

1) state/county or CBSA 
    * Name of the county if it is a county.


2) year 
    * Value of year


3) counties_msa 
    * Names of all counties belonging to the Metro Area if it is a Metro Area
    (MSA).


4) town_name 
    * Town name - applicable for North East regions


5) metro_status 
    * value will be "1" if it is a metropolitan county. Otherwise value will be
    "0".


6) metro_name 
    * Metro area name if metro_status is "1"


7) smallarea_status 
    * value will be "1" if it is a small area. Otherwise value will be "0".


8) Efficiency(A studio apartment) 
    * Efficiency FMR in US Dollars


9) One-Bedroom 
    * 1-bedroom FMR in US Dollars


10) Two-Bedroom 
    * 2-bedroom FMR in US Dollars


11) Three-Bedroom 
    * 3-bedroom FMR in US Dollars


12) Four-Bedroom 
    * 4-bedroom FMR in US Dollars
    

### References

"What Is Fair Market Rent? | RentData.org." Www.rentdata.org,
www.rentdata.org/articles/what-is-fair-market-rent. Accessed 18 Feb. 2022.

"Fair Market Rents | HUD USER." Huduser.gov, U.S Department of Housing and Urban
Development, 2017, www.huduser.gov/portal/datasets/fmr.html.
