---
title: "Income-Limits"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Income-Limits}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Program Overview

"The Department of Housing and Urban Development (HUD) sets income limits that
determine eligibility for assisted housing programs including the Public
Housing, Section 8 project-based, Section 8 Housing Choice Voucher, Section 202
housing for the elderly, and Section 811 housing for persons with disabilities
programs. HUD develops income limits based on Median Family Income estimates and
Fair Market Rent area definitions for each metropolitan area, parts of some
metropolitan areas, and each non-metropolitan
county("Income Limits | HUD USER")."

### Example

This is a basic example showing you how to query the Income Limits API.

Querying a state will return a data frame with rows representing counties 
within the state. A (state fip + county fip + county subdivision) 10 digit code
will return a 1 row data frame for that county. A CBSA code will also return a 1 
row data frame for that area.

This works similarly to the FMR API.

```{r, eval=FALSE}
library(hudr)

# A list of states and some associated metadata can be provided by:
hud_states()

# This gets FMR data for the state of Virginia in 2021.
hud_il(query = 'VA', year = '2021')

# A list of counties and associated metadata can be provided by:
# This queries counties for Virginia.
hud_counties("VA")

# Getting a county requires a state fip + county fip + 99999
hud_il(query = '0100199999', year = '2017')

# You can also check for CBSA codes which can also be supplied in hud_fmr()
# query field.
hud_il("METRO47900M47900", year=c(2018))

# A list of metropolitan areas(CBSA) and associated metadata can be provided by:
hud_metropolitan()
```

### Input Arguments

**QUERY**: Can provide either a 10 digit FIPS code which is almost always
  state fips + county fips + 99999, or state abbreviation. 
  Can also provide a CBSA code. This CBSA code requires both itself and its
  corresponding division code. You are only allowed to query for metropolitan areas.
  
  Run hud_states() to get a list of counties
  
  Run hud_metropolitan() to get a list of metropolitan areas.
  
  Run hud_counties("MD") to get list of counties in MD
  
  * query = '5100199999'
  * query = 'MD'
  * query = 'METRO12700M12700'

**YEAR**: Year range of the data to retrieve. Defaults to current year. Make sure 
the years provided follow the rules provided by the TYPE argument.

  * year = c(2019, 2018, 2021)
  * year = c(2016, 2017, 2022)
  * year = c(2021)

**KEY**: The API key.

  * key = "wqokqo2138jdi13wfwwfwcytjyr"

### Output Data

1) **county_name** 
    * Name of the county if it is a county.


2) **counties_msa**
    * Names of all counties belonging to the Metro Area if it is a Metro Area (MSA).


3) **town_name**
    * Town name - applicable for North East regions


4) **metro_status**
    * value will be "1" if it is a metropolitan county. Otherwise value will be "0".


5) **metro_name**
    * Metro area name if metro_status is "1"


6) **year**
    * year of the data


7) **median_income**
    * Median Income for the area
    
    
8) **Very Low** 
    * (50%) Income Limits
    * For more details on il50 check the website.


9) **Extremely Low** 
    * (30%) Income Limits
    * For more details on il30 check the website.
    
    
10) **Low** 
    * (80%) Income Limits
    * For more details on il80 check the website.
    
### References

"Income Limits | HUD USER." Huduser.gov, 2015,
www.huduser.gov/portal/datasets/il.html.
