---
title: "Income-Limits"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Income-Limits}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Program Overview

According to (Income Limits | HUD USER), HUD establishes income restrictions for
assisted housing programs such as Public Housing, Section 8 project-based,
Section 8 Housing Choice Voucher, Section 202 housing for the elderly, and
Section 811 housing for persons with disabilities. For each metropolitan region,
parts of some metropolitan areas, and non-metropolitan county, HUD creates
income limitations based on Median Family Income estimates and Fair Market Rent
area criteria.

### Examples

This page contains basic examples of how to query the HUD Income Limits API.
Before looking at the outputted data I RECOMMEND first taking a look at [Input Arguments][] as well as [Output Data][] located at the bottom of the page.

**DISCLAIMER:** The output tables are too large to be displayed here so we
choose to leave them out.

* Querying a state will return a single row containing the Income Limits for that
state. 
* Querying for a county will return the IL measurement for that county.
* Querying for a cbsa will return the measurement for that cbsa.

#### US States
This is an example to show you how to query for all states and territories
in the US.
```{r, eval=FALSE}
library(rhud)
hud_nation_states_territories()
```

#### US Counties 
This is an example to show you how to query for all counties in MD,
Virginia, and California, respectively.
```{r, eval=FALSE}
hud_state_counties("MD")

hud_state_counties("Virginia")

hud_state_counties("6")
```

#### US Metropolitan Areas
This is an example to show you how to query for metropolitan areas for a state.
```{r, eval=FALSE}
# Get all metropolitan areas in Maryland and Virginia.
hud_state_metropolitan(c("MD", "VA"))
```

#### State Level Income Limits
These are examples which show you how to query the Income Limits on a state
level basis. You can use the state abbreviation, state name, or state fipscode.
It will return a single row indicating the income limits for that particular
state.

```{r, eval=FALSE}
# These function calls gets income limits data for the state of Virginia in 2021.
hud_il(query = 'VA', year = '2021')

hud_il(query= "Virginia", year = '2021')

hud_il(query= "51", year = '2021')

# You can also choose to query for multiple states or multiple years.
hud_il(query = c('VA', 'MD'), year = c('2021', '2019'))
```

#### County Level Income Limits
These are examples that show you how to query the Income Limits on a 
county-level basis: counties are defined with a 2 digit state fipscode + 3 digit county fipscode. You also need a 99999 code added to the end. We currently 
don't know why the extra 99999 is included and have not found any cases where it's different. I recommend looking at [US Counties][] to determine what counties 
are available for particular states.

```{r, eval=FALSE}
# Getting a county requires a 2 digit state fip + 3 digit county fip + 99999
# This queries for Income Limits in 2017.
hud_il(query = '0100199999', year = '2017')
```

#### Small Areas Income Limits
These are examples that show you how to query the Small Areas Income Limits.
Small Areas are defined as metropolitan and micropolitan areas specified using
CBSA codes. Currently, we know how to query for metropolitan small areas, but
are not sure for micropolitan yet. I recommend looking at 
[US Metropolitan Areas][] to determine what counties are available.
```{r, eval=FALSE}
# Gets FMR data for METRO area.
hud_il(query = "METRO47900M47900", year=c(2018))
```

### Input Arguments

**QUERY**: Can provide either a 10 digit FIPS code which is almost always state
fips + county fips + 99999, or state abbreviation. Can also provide a CBSA code.
Currently, you are only allowed to query for metropolitan areas.
  
  Run hud_states() to get a list of states in the US.
  
  Run hud_metropolitan("MD") to get a list of metropolitan areas in MD.
  
  Run hud_counties("MD") to get list of counties in MD.
  
  * query = '5100199999'
  * query = 'MD'
  * query = 'METRO12700M12700'

**YEAR**: Year range of the data to retrieve: defaults to current year. 

  * year = c(2019, 2018, 2021)
  * year = c(2016, 2017, 2022)
  * year = c(2021)

**KEY**: The API key.

  * key = "wqokqo2138jdi13wfwwfwcytjyr"

### Output Data

1) **county_name** 
    * Name of the county if it is a county.


2) **counties_msa**
    * Names of all counties belonging to the Metro Area if it is a Metro Area (MSA).


3) **town_name**
    * Town name - applicable for North East regions


4) **metro_status**
    * value will be "1" if it is a metropolitan county. Otherwise value will be "0".


5) **metro_name**
    * Metro area name if metro_status is "1"


6) **year**
    * year of the data


7) **median_income**
    * Median Income for the area
    
    
8) **Very Low** 
    * (50%) Income Limits
    * For more details on il50 check the website.


9) **Extremely Low** 
    * (30%) Income Limits
    * For more details on il30 check the website.
    
    
10) **Low** 
    * (80%) Income Limits
    * For more details on il80 check the website.
    
### References

"Income Limits | HUD USER." Huduser.gov, 2015,
<br />&nbsp;&nbsp;&nbsp;&nbsp; www.huduser.gov/portal/datasets/il.html.
