---
title: "Setup"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Setup}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Setup

You can install the development version from GitHub with:

```{r, warning=FALSE, results='hide',message=FALSE}
# install.packages("devtools")
library(devtools)
devtools::install_github("etam4260/hudr")
```

The US Department of Housing and Urban Development for HUD USER requires users 
to gain an  access key before querying their systems. You must go to https://www.huduser.gov/hudapi/public/register?comingfrom=1 and follow the 
instructions for making an account.

HUD Exchange has data stored in files and does not have a readily available API
system for gaining access to their data. 

From there you need to make a new token. Make sure to save the token somewhere 
as you will only be able to view it once. However, you can use it as many times 
as you want. You can now supply the 'key' argument to the function calls. 
However, it is **NOT RECOMMENDED** to do this. Rather, to reduce the need to 
supply the key argument in every function call, you can use the Sys.setenv() 
function to save the key to the user environment.

## Key Setup

There are various method of setting up a key. However, I recommend using THE
Sys.setenv() function. You can also choose to save it into your .Rprofile or
.Renviron files. 

### Using Sys.setenv() during an interactive session
One method of setting the API key is to use environment variables. This library
searches for HUD_KEY in the user environment to check if a key has been defined.
If no key has been defined, an empty string is outputted.

```{r, eval=FALSE}
library(hudr)

Sys.getenv("HUD_KEY")
```

To set the key...

```{r, eval=TRUE}
Sys.setenv("HUD_KEY" = "2yeuduhq72ueajk")
Sys.getenv("HUD_KEY")
```

To check whether hudr can gain access to this environment variable...

```{r, eval=TRUE}
hudgetkey()
```

It is now setup for the rest of the R session.

You can now access the API from HUD User. I recommend taking a look at the other
articles to get a better sense of how to access specific datasets. For more 
details on what datasets are available and their official documentation, go to
https://www.huduser.gov/portal/datasets/update-schedule.html

You can also check available datasets shown at 
https://www.huduser.gov/portal/datasets/update-schedule.html and those
which this library currently supports.

```{r, eval=TRUE}
huddatasets()
```
