---
title: "USPS-Crosswalk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{USPS-Crosswalk}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Program Overview

According to (US Department of Housing and Urban Development, n.d.), the
difficulty of linking United States Postal Service (USPS) ZIP codes to Census
Bureau regions is one of the many obstacles that social science scholars and
practitioners face. Only at the ZIP code level is there relevant data that, when
paired with demographic data calculated at various Census geographic levels,
could open up new paths of investigation.

Furthermore, the (US Department of Housing and Urban Development, n.d.) believes
that while there are several appropriate approaches for integrating ZIP codes
with Census geographies, they are limited. The HUD-USPS Crosswalk Files were
supplied by PD&R to enable more routes for integrating these data. These
one-of-a-kind files were created using data from the quarterly USPS Vacancy
Data. They come straight from the USPS; they're updated quarterly, so they're
always up to date with changes in ZIP code configurations; and they reflect both
commercial and residential addresses. Because many of the phenomena that housing
researchers study are based on housing unit or address, the latter attribute is
of special interest to them. Analysts can take into account not just the spatial
distribution of population, but also the spatial distribution of residences, by
utilizing an allocation approach based on residential addresses rather than by
area or population. This allows for a more sophisticated approach to data
allocation across different geographies.

These journal articles describe the problem and proposed solution in more detail: 

|
| Wilson, Ron and Din, Alexander, 2018. “Understanding and Enhancing the U.S.
|         Department of Housing and Urban Development’s ZIP Code Crosswalk Files,”
|         Cityscape: A Journal of Policy Development and Research, Volume 20 Number 2, 277
|         https://www.huduser.gov/portal/periodicals/cityscpe/vol20num2/ch16.pdf
|
|
| Din, Alexander and Wilson, Ron, 2020. "Crosswalking ZIP Codes to Census
|         Geographies: Geoprocessing the U.S. Department of Housing & Urban Development’s
|         ZIP Code Crosswalk Files," Cityscape: A Journal of Policy Development and
|         Research, Volume 22, Number 1,
|         https://www.huduser.gov/portal/periodicals/cityscpe/vol22num1/ch12.pdf

<br>

## Census Geographies

This chart provided by the US Census Bureau gives a good example of the
relationships among different geographies. The crosswalk files only support a
subset of these.

<br>

## Examples

There are 12 main function calls for the crosswalk files:
the package also contains an omni function which encapsulates the 
capabilities of all the main function calls below --
[omni-function][Using the omni function for querying]

1) [zip-tract][Crosswalk zipcode to census tract]
2) [zip-county][Crosswalk zipcode to county fip]
3) [zip-cbsa (Core Base Statistical Areas)][Crosswalk zipcode to core base statistical area (cbsa)]
4) [zip-cbsadiv (Available 4th Quarter 2017 onwards)][Crosswalk zipcode to core based statistical area division (cbsadiv)]
5) [zip-cd (Congressional District)][Crosswalk zipcode to congressional district (cd)]
6) [tract-zip][Crosswalk census tract to zipcode]
7) [county-zip][Crosswalk county fip into zipcode]
8) [cbsa-zip][Crosswalk core based statistical areas (cbsa) to zipcode]
9) [cbsadiv-zip (Available 4th Quarter 2017 onwards)][Crosswalk core based statistical areas division (cbsadiv) to zipcode]
10) [cd-zip][Crosswalk congressional district (cd) to zipcode]
11) [zip-countysub (Available 2nd Quarter 2018 onwards)][Crosswalk zipcode to county subdivision (countysub)]
12) [countysub-zip (Available 2nd Quarter 2018 onwards)][Crosswalk county subdivision (countysub) to zipcode]

The first geoid type in the function call is what to query for. For
example in 1) above, 'zip' is the first geoid and 'tract' is the second geoid.

The second geoid in the function call describes the geoid which we want to 
determine 'intersection' with the first geoid where intersection is described
as the % of residential, business, other, and total buildings that overlap. 

For example, in function call #7, we might have a county called 22031 
which has zip codes 71052, 71078, 71049, 71032 ... where the residential 
% (res_ratio) of each zip is 0.38, 0.21, 0.11, 0.05 ... respectively. 
Of all these zipcodes' res_ratios, when added up will equal 1,
signaling these grouping of zip codes make up 100% of residential address in the
county with each zipcode taking up their respective residential percentage.

Disclaimer: Although there exists inverse relationships in the Crosswalk 
Files, the measurements are NOT COMPLETELY inverse -- for reasons stated
within the papers above.

These are basic examples which shows you to query the Crosswalk API. Before
looking at the outputted data I RECOMMEND first taking a look at [the parameters][Parameters]
as well as [return data][Returns] located at the bottom of the page.

### Crosswalk zipcode to census tract
```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE}
library(rhud)
options (digits=4)

hud_cw_zip_tract(zip = '35213', year = c('2010'), quarter = c('1'))
```
<br>

### Crosswalk zipcode to county fip
```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE}
hud_cw_zip_county(zip = 35213, year = c('2020'), quarter = c('2'))
```
<br>

### Crosswalk zipcode to core base statistical area (cbsa)
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_zip_cbsa(zip = 35213, year = c('2011'), quarter = c('3'))
```
<br>

### Crosswalk zipcode to core based statistical area division (cbsadiv)
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_zip_cbsadiv(zip = '22031', year = c('2019'), quarter = c('4'))
```
<br>

### Crosswalk zipcode to congressional district (cd)
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_zip_cd(zip = '35213', year = c(2011), quarter = c(1))
```
<br>

### Crosswalk census tract to zipcode
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_tract_zip(tract = 48201223100, year = c('2017'), quarter = c('1'))
```
<br>

### Crosswalk county fip into zipcode 
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_county_zip(county = '22031', year = c('2010'), quarter = c('1'))
```
<br>

### Crosswalk core based statistical areas (cbsa) to zipcode
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_cbsa_zip(cbsa = '10140', year = c('2017'), quarter = c('2'))
```
<br>

### Crosswalk core based statistical areas division (cbsadiv) to zipcode
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_cbsadiv_zip(cbsadiv = 10380, year = c('2017'), quarter = c('4'))
```
<br>

### Crosswalk congressional district (cd) to zipcode
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_cd_zip(cd = '2202', year = c('2010'), quarter = c('4'))
```
<br>

### Crosswalk zipcode to county subdivision (countysub)
```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE}
hud_cw_zip_countysub(zip = '35213', year = c('2019'), quarter = c('2'))
```
<br>

### Crosswalk county subdivision (countysub) to zipcode 
```{r, eval = if(Sys.getenv("HUD_KEY") == "")  FALSE else TRUE}
hud_cw_countysub_zip(countysub = '4606720300 ', year = c('2019', '2019', '2019'),
                     quarter = c('4','4'))
```

<br>

## Querying for only the crosswalked geoids

If you just want the crosswalked geoids, you can set the **minimal** argument 
to TRUE. This will return a vector containing the crosswalked geoids
without the extra metadata.

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE }
hud_cw_county_zip(county = '22031', year = c('2010'), quarter = c('1'), minimal = TRUE)
```

<br>

## Using the omni function for querying

The omni function is a redundant implementation of the functions shown above
that requires specifying the type which can be from 1-12. The type argument
follows the number scheme described at within the [input arguments][Input Arguments]. You also 
must use the 'query' argument (i.e query = 22031) for inputting geoids instead 
of the specific geoid names (i.e county = 22031, cd = 7200) used by the 
above functions.

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE }
hud_cw(type = 7, query = '22031', year = c('2010'), quarter = c('1'))
```

<br>

## Crosswalking a dataset

For those who need to apply an allocation method (residential, business, other,
total) to individual items in a data set, the **crosswalk()** function is available. 
Lets say we wanted to know the population at a zip code level (there is likely 
already a data set for this) for the counties of Washington, Wicomico, 
and Worchester in Maryland in the year 2019. 

NOTE: The use of the crosswalk() function is likely best suited for datasets 
that are not described in the geographic identifier we want to crosswalk to. In
this case population might not be the best example for this.

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE }
sample <- data.frame(pop = c(151049, 103609, 52276),
                     county = c("24043", "24045", "24047"))

head(sample)
```

In the crosswalked data set below each zip code associated with a county 
is assigned the same population value.

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE }
crosswalk(data = sample, geoid = "county", geoid_col = "county", 
          cw_geoid = "zip", cw_geoid_col = NA, method = NA, year = 2019
          , quarter = 1)
```

<br>

To utilize an allocation method provided by the crosswalk files and apply it
to columns of the data set, specify the method and cw_geoid_col arguments.
In this case we want to allocate the county population levels to a zip code
level using the method based on the ratio of residential addresses.

<br>

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE}
crosswalk(data = sample, geoid = "county", geoid_col = "county", 
          cw_geoid = "zip", cw_geoid_col = "pop", method = "res", year = 2019
          , quarter = 1)
```
<br>

## Geo-in-geo functions: does one geography overlap another?

This library also allows the user to determine if one geography overlaps
another.

<b>Disclaimer:</b> Overlap is not determined by whether the boundaries intersect, but
rather if any addresses lie in both. 

The function below queries for if zipcode 71052 overlaps cbsa 43340. 

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE }
71052 %z_in_cbsa% 43340

```
<br>

The infix operator used above defaults to the most recent crosswalk files 
available. Furthermore, you must have the HUD_KEY system environment variable 
set before using them: check the Setup page.

The non infix version is also available: this allows you to specify the year,
quarter, and key arguments.

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE }
z_in_cbsa(zip = 71052, cbsa = 43340, year = 2019, quarter = 2)

```
<br>

## Most recent USPS Crosswalk files

To get the most recent crosswalk files by year and quarter:

```{r, eval = if(Sys.getenv("HUD_KEY") == "") FALSE else TRUE }
hud_rec_cw_yr()

```

<br><br><br>

## Parameters

<table>
	<tr>
		  <th>Parameter</th>
		  <th>Type</th>
		  <th>Description</th>
 	</tr>
	<tr>
  		<td>query</td>
   		<td>
          The query argument name is the first geoid in the function call. 
          For example, zip-county would have the 'zip' argument name as the query.  
          <br>
          Functions #1-5 and #11 must specify a 5 digit zip code in the **zip** function
          argument. 
          <br><br>
          * zip = 22031
          
          <br>
          Function #6 must specify an 11 digit number consisting of state FIPS + county
          FIPS + tract code tract in the **tract** function argument.
          <br><br>
          * tract = 51059461700
          
          <br>
          Function #7 must specify a 5 digit county fips code in the **county** function
          argument.
          <br><br>
          * county = 51600
          
          <br>
          Function #8 must specify a micropolitan or metropolitan CBSA code in the
          **cbsa** function argument. 
          <br><br>
          * cbsa = 10380
          
          <br>
          Function #9 must specify a metropolitan CBSA division code in the
          **cbsadiv** function argument. 
          <br><br>
          * cbsadiv = 35614
          
          <br>
          I recommend typing hud_metropolitan("MD") to see a list of metropolitan CBSA in
          Maryland. For example, a code of METRO22900N40079 needs 22900 inside as the input
          argument. 
          <br><br>
          Function #10 must be a 2 digit state fips code + 2 digit congressional district 
          code in the **cd** function argument. 
          <br><br>
          * cd = 7200
          
          <br>
          Function #12 must be a 10 digit number consisting of 5 digit county fips code + 5 digit
          county sub district code in the **countysub** argument. 
          <br><br>
          * countysub = 4606720300
          
   		    <br>
   		</td>
 	</tr>
	<tr>
  		<td>year</td>
   		<td>
          Years of the data to retrieve: defaults to the current year.
          <br><br>
          * year = c(2019, 2018, 2021)
          * year = c(2016)
          * year = 2021
   		</td>
 	</tr>
	<tr>
  		<td>quarter</td>
   		<td>
         The quarters in the year to retrieve: defaults to the first
         quarter. 
         <br><br>
         * quarter = c(1,2,3,4)
         * quarter = c(1)
         * quarter = 4
   		</td>
 	</tr>
 	<tr>
  		<td>minimal</td>
   		<td>
         If TRUE, returns just the intersecting geoids that are 
         crosswalked with the queried geoid.  
   		</td>   
 	</tr>
  <tr>
  		<td>key</td>
   		<td>
         The API key provided by HUD USER.
         <br><br>
         * key = "wqokqo2138jdi13wfwwfwcytjyr"
   		</td>   
 	</tr>
</table>

<br><br><br>

## Returns

<table>
	<tr>
		  <th>Return Field</th>
		  <th>Description</th>
 	</tr>
 	<tr>
  		<td>zip,<br>county,<br>fips,<br>tract,<br>cbsa,<br>cbsadiv,<br>cd,<br>countysub</td>
   		<td>The geoid that is queried for. The name of the column will be named the geoid type</td>
 	</tr>
	<tr>
  		<td>res_ratio</td>
   		<td> The ratio of residential addresses in the ZIP – Tract, County, or CBSA
      part to the total number of residential addresses in the entire ZIP. (for
      type 1-5 and 11) The ratio of residential addresses in the Zip, Tract, County, or
      CBSA - ZIP part to the total number of residential addresses in the entire
      Zip, Tract, County, or CBSA. (for type 6-10 and 12)</td>
 	</tr>
	<tr>
  		<td>bus_ratio</td>
   		<td>The ratio of business addresses in the ZIP – Tract, County, or CBSA part
      to the total number of business addresses in the entire ZIP. (for type 1-5 and 11).
      The ratio of business addresses in the Tract, County, or CBSA – ZIP part to
      the total number of business addresses in the entire Tract, County, or CBSA.
      (for type 6-10 and 12)</td>
 	</tr>
	<tr>
  		<td>oth_ratio</td>
   		<td>The ratio of other addresses in the ZIP – Tract to the total number of
      other addresses in the entire ZIP. (for type 1-5 and 11). The ratio of other 
      addresses in the Tract, County, or CBSA – ZIP part to the total number of
      other addresses in the entire Tract, County, or CBSA. (for type 6-10 and 12)</td>
 	</tr>
 	<tr>
  		<td>tot_ratio</td>
   		<td>The ratio of all addresses in the ZIP – Tract to the total number of 
      all types of addresses in the entire ZIP. (for type 1-5 and 11) The ratio of all
      addresses in the Tract, County, or CBSA-ZIP part to the total number of all
      types of addresses in the entire Tract, County, or CBSA. (for type 6-10 and 12)
      </td>   
 	</tr>
 	<tr>
  		<td>zip,<br>county,<br>fips,<br>tract,<br>cbsa,<br>cbsadiv,<br>cd,<br>countysub</td>
   		<td>The intersecting geoids depending on function call</td>   
 	</tr>
 	<tr>
  		<td>year</td>
   		<td>Year the measurement was taken.</td>   
 	</tr>
 	<tr>
  		<td>quarter</td>
   		<td>Quarter of year when measurement was taken.</td>   
 	</tr>
</table>
		

<br>

## References

| Din, Alexander and Wilson, Ron, 2020. "Crosswalking ZIP Codes to Census 
|       Geographies: Geoprocessing the U.S. Department of Housing & Urban Development’s
|       ZIP Code Crosswalk Files," Cityscape: A Journal of Policy Development and
|       Research, Volume 22, Number 1, https://www.huduser.gov/portal/periodicals/cityscpe/vol22num1/ch12.pdf
|   
| Katy Rossiter, K. R. (2014, July 31). Standard Hierarchy of Census Bereau
|       Geographies [Photograph]. Understanding Geographic Relationships: Counties,
|       Places, Tracts and More.
|       https://www.census.gov/newsroom/blogs/random-samplings/2014/07/understanding-geographic-relation
|       ships-counties-places-tracts-and-more.html
|
| U.S Department of Housing and Urban Development. (n.d.). HUD USPS ZIP
|       Code Crosswalk Files | HUD USER. HUD  USPS ZIP CODE CROSSWALK FILES.
|       Retrieved February 17, 2022, from 
|       https://www.huduser.gov/portal/datasets/usps_crosswalk.html
|
| Wilson, Ron and Din, Alexander, 2018. “Understanding and Enhancing the U.S.
|       Department of Housing and Urban Development’s ZIP Code Crosswalk Files,”
|       Cityscape: A Journal of Policy Development and Research, Volume 20 Number 2, 277
|       – 294.
