---
title: "Visualizing Fair Market Rents and Income Limits"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing-FMR-and-IL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<i>Sample provided by <b>Shuyu Jin (geojsy@umd.edu)</b></i>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

### Overview
The US Department of Housing and Urban Development maintains both [Fair Market Rents](https://www.hud.gov/program_offices/public_indian_housing/programs/hcv/landlord/fmr) (FMR) and [Income Limits](https://www.huduser.gov/portal/datasets/il.html) (IL) programs which are crucial for determining eligibility for assistance dollars in the US. Provided below is an example of querying both FMR and IL as well as mapping them using the leaflet R library.


#### Loading in the necessary libraries

```{r message=FALSE}
library(rhud)
library(tigris)
library(leaflet)
library(tidyverse)
library(sf)
```

To improve performance, use tigris caching.

```{r message=FALSE}
options(tigris_use_cache = TRUE)
```

#### Reading in data

Read in the Fair Markets Rent for counties in the year 2021 as well as the shape
files needed to map.

```{r message=FALSE}
md_counties_2021_fmr <- hud_fmr_state_counties(state = 'MD', year = '2021') 
```

Get county data for the state of Maryland.

```{r message=FALSE}
md_counties <- hud_state_counties("MD")
```

Query for every county and get their Income Limits data using the rhud package.

```{r message=FALSE}
md_counties_2021_il <- hud_il(md_counties$fips_code)

#Find unique column names and keep only unique column names
unique_names <- unique(colnames(md_counties_2021_il))
md_counties_2021_il_unique <- md_counties_2021_il[unique_names]
```

Download the shape files for counties in the US and obtain consistent projection.

```{r message=FALSE}
county_shapefile <- tigris::counties(state = NULL, cb = FALSE,
                    resolution = "500k", year = 2021) %>% 
                    sf::st_transform('+proj=longlat +datum=WGS84')

```

Filter the counties to those only in Maryland.

```{r message=FALSE}
md_county_shapefile <- county_shapefile[county_shapefile$STATEFP == 24, ]

md_county_shapefile$fullFips <- paste0(as.character(md_county_shapefile$STATEFP),
                                 as.character(md_county_shapefile$COUNTYFP),
                                 "99999")
```


#### Merging of shapefiles and data

Merge the shape file with the FMR and IL data.

```{r message=FALSE, warning=FALSE}
md_county_fmr_merged <- tigris::geo_join(md_county_shapefile,
                                        md_counties_2021_fmr,
                                        by_sp = "fullFips",
                                        by_df = "fips_code",
                                        how = "inner")
```

```{r message=FALSE, warning=FALSE}
md_county_il_merged <- tigris::geo_join(md_county_shapefile,
                                       md_counties_2021_il_unique,
                                       by_sp = "fullFips",
                                       by_df = "query",
                                       how = "inner")
```


### Leaflet Mapping

Create the pallete for coloring the overlayed shape file. The scale used
corresponds to Efficiency data and Median Income data.

```{r message=FALSE}
pal_md_fmr <- colorNumeric(
  palette = "Greens",
  domain = as.numeric(md_county_fmr_merged$Efficiency))
```

```{r message=FALSE}
pal_md_il <- colorNumeric(
  palette = "Purples",
  domain = as.numeric(md_county_il_merged$median_income))
```


Create the county level, efficiency, one bedroom, two bedroom, three bedroom
four bedroom label to appear when hovering using the FMR overlay.


```{r message=FALSE}
label_fmr <- paste0(
  "County: ", md_county_fmr_merged$NAME, "<br/>",
  "Efficiency: ", scales::dollar(as.numeric(md_county_fmr_merged$Efficiency)), "<br/>",
  "One Bedroom: ", scales::dollar(as.numeric(md_county_fmr_merged$`One-Bedroom`)), "<br/>",
  "Two Bedroom: ", scales::dollar(as.numeric(md_county_fmr_merged$`Two-Bedroom`)), "<br/>",
  "Three Bedroom: ", scales::dollar(as.numeric(md_county_fmr_merged$`Three-Bedroom`)), "<br/>",
  "Four Bedroom: ", scales::dollar(as.numeric(md_county_fmr_merged$`Four-Bedroom`))) %>% 
  lapply(htmltools::HTML)
```

Create the county level, median income, 1 person family low/very low/extremely low income limits label to appear when hovering using the IL overlay

```{r message=FALSE}
label_il_p1 <- paste0(
  "County: ", md_county_il_merged$NAME, "<br/>",
  "Median Income: ", scales::dollar(as.numeric(md_county_il_merged$median_income)), "<br/>",
  "1 Person Family Low (80%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$low.il80_p1)), "<br/>",
  "1 Person Family Very Low (50%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$very_low.il50_p1)), "<br/>",
  "1 Person Family Extremely Low (30%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$extremely_low.il30_p1))) %>% 
  lapply(htmltools::HTML)
```


Create the county level, median income, 2 person family low/very low/extremely low income limits label to appear when hovering using the IL overlay

```{r message=FALSE}
label_il_p2 <- paste0(
  "County: ", md_county_il_merged$NAME, "<br/>",
  "Median Income: ", scales::dollar(as.numeric(md_county_il_merged$median_income)), "<br/>",
  "2 Person Family Low (80%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$low.il80_p2)), "<br/>",
  "2 Person Family Very Low (50%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$very_low.il50_p2)), "<br/>",
  "2 Person Family Extremely Low (30%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$extremely_low.il30_p2))) %>% 
  lapply(htmltools::HTML)
```


Create the county level, median income, 3 person family low/very low/extremely low income limits label to appear when hovering using the IL overlay

```{r message=FALSE}
label_il_p3 <- paste0(
  "County: ", md_county_il_merged$NAME, "<br/>",
  "Median Income: ", scales::dollar(as.numeric(md_county_il_merged$median_income)), "<br/>",
  "3 Person Family Low (80%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$low.il80_p3)), "<br/>",
  "3 Person Family Very Low (50%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$very_low.il50_p3)), "<br/>",
  "3 Person Family Extremely Low (30%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$extremely_low.il30_p3))) %>% 
  lapply(htmltools::HTML)
```


Create the county level, median income, 4 person family low/very low/extremely low income limits label to appear when hovering using the IL overlay

```{r message=FALSE}
label_il_p4 <- paste0(
  "County: ", md_county_il_merged$NAME, "<br/>",
  "Median Income: ", scales::dollar(as.numeric(md_county_il_merged$median_income)), "<br/>",
  "4 Person Family Low (80%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$low.il80_p4)), "<br/>",
  "4 Person Family Very Low (50%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$very_low.il50_p4)), "<br/>",
  "4 Person Family Extremely Low (30%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$extremely_low.il30_p4))) %>% 
  lapply(htmltools::HTML)
```


Create the county level, median income, 5 person family low/very low/extremely low income limits label to appear when hovering using the IL overlay

```{r message=FALSE}
label_il_p5 <- paste0(
  "County: ", md_county_il_merged$NAME, "<br/>",
  "Median Income: ", scales::dollar(as.numeric(md_county_il_merged$median_income)), "<br/>",
  "5 Person Family Low (80%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$low.il80_p5)), "<br/>",
  "5 Person Family Very Low (50%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$very_low.il50_p5)), "<br/>",
  "5 Person Family Extremely Low (30%) Income Limits: ", scales::dollar(as.numeric(md_county_il_merged$extremely_low.il30_p5))) %>% 
  lapply(htmltools::HTML)
```

Using the leaflet package, plot the county level map of Fair Markets Rents and 
Income Limits in the state of Maryland.

```{r}
# The legends overlap with the zoom control buttons, need to move these controls.

leaflet(width = "100%", options = leafletOptions(zoomControl = FALSE)) %>%
  htmlwidgets::onRender("function(el, x){
                        L.control.zoom({position: 'topright'}).addTo(this)
                        }")%>%
  
  addProviderTiles("CartoDB") %>% 
  
  addPolygons(data = md_county_fmr_merged, fillColor = ~pal_md_fmr(as.numeric(Efficiency)),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(weight = 2,
                                           color = "#666",
                                           dashArray = "",
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),
              label = label_fmr,
              group = "Fair Markets Rent") %>%
  
  addPolygons(data = md_county_il_merged, fillColor = ~pal_md_il(as.numeric(median_income)),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(weight = 2,
                                           color = "#666",
                                           dashArray = "",
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),
              label = label_il_p1,
              group = "Income Limits-1 person") %>% 
  
  addPolygons(data = md_county_il_merged, fillColor = ~pal_md_il(as.numeric(median_income)),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(weight = 2,
                                           color = "#666",
                                           dashArray = "",
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),
              label = label_il_p2,
              group = "Income Limits-2 person") %>%   
  
    addPolygons(data = md_county_il_merged, fillColor = ~pal_md_il(as.numeric(median_income)),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(weight = 2,
                                           color = "#666",
                                           dashArray = "",
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),
              label = label_il_p3,
              group = "Income Limits-3 person") %>%   
  
    addPolygons(data = md_county_il_merged, fillColor = ~pal_md_il(as.numeric(median_income)),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(weight = 2,
                                           color = "#666",
                                           dashArray = "",
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),
              label = label_il_p4,
              group = "Income Limits-4 person") %>%   
  
    addPolygons(data = md_county_il_merged, fillColor = ~pal_md_il(as.numeric(median_income)),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(weight = 2,
                                           color = "#666",
                                           dashArray = "",
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),
              label = label_il_p5,
              group = "Income Limits-5 person") %>%   
  
  addLayersControl(overlayGroups = c("Fair Markets Rent", "Income Limits-1 person", "Income Limits-2 person", "Income Limits-3 person", "Income Limits-4 person","Income Limits-5 person"), baseGroups = "CartoDB") %>%
  
  # add legend
  addLegend("bottomright",
    pal = pal_md_fmr,
    values = as.numeric(md_county_fmr_merged$Efficiency),
            opacity = 0.7,
            title = htmltools::HTML("Fair Markets Rent <br>
                                    by County"), group = "Fair Markets Rent") %>%
  
    addLegend("bottomleft",
      pal = pal_md_il,
            values = as.numeric(md_county_il_merged$median_income),
            opacity = 0.7,
            title = htmltools::HTML("Income Limits <br>
                                    by County"), group = "Income Limits") %>%
setView(lat = 39, lng = -77, zoom = 6)
```



### Observations
Using the rhud package to download FMR and IL data, combined with the Leaflet package for map visualization, users can understand the data more intuitively.



### Reference
U.S. Department of Housing and Urban Development. Fair Market Rents (40th Percentile Rents). https://www.huduser.gov/portal/datasets/fmr.html. (Accessed 04/03/22)

U.S. Department of Housing and Urban Development. Income Limits. https://www.huduser.gov/portal/datasets/il.html. (Accessed 04/03/22)


