---
title: "Visualizing-FMR-and-IL"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Visualizing-FMR-and-IL}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

<i>Sample provided by <b>Shuyu Jin</b> (geojsy@umd.edu)</i>

## Overview

The US Department of Housing and Urban Development maintains both
[FMR](https://link-url-here.org) (Fair Markets Rent) and 
[IL](https://link-url-here.org) (Income Limits) programs which are crucial
for determining eligibility for assistance dollars in the US. Provided below is
an example of querying both FMR and IL as well as mapping them using 
the leaflet R library.

## Loading in the necessary libraries

```{r message=FALSE}
library(rhud)
library(tigris)
library(leaflet)
library(tidyverse)
library(sf)
```

<br>

To improve performance, use tigris caching.

```{r message=FALSE}
options(tigris_use_cache = TRUE)
```

## Reading in data

Read in the Fair Markets Rent for counties in the year 2021 as well as the shape
files needed to map.

```{r message=FALSE}
md_counties_2021_fmr = hud_fmr_state_counties(state = 'MD', year = '2021')
```

<br>

Get county data for the state of Maryland using the rhud package.

```{r message=FALSE}
md_counties = hud_state_counties("MD")
```

<br>

Query for every county in Maryland and get their Income Limits data using
the rhud package.

```{r message=FALSE}
md_counties_2021_il = hud_il(md_counties$fips_code)
```

<br>

Download the shape files for counties in the US.

```{r message=FALSE}
md_county_shapefile = tigris::counties(state = NULL, cb = FALSE,
                                       resolution = "500k", year = 2021)
```

<br>

Filter the counties to those only in Maryland.

```{r message=FALSE}
md_county_shapefile = md_county_shapefile[md_county_shapefile$STATEFP == 24, ]

md_county_shapefile$fullFips = paste0(as.character(md_county_shapefile$STATEFP),
                                 as.character(md_county_shapefile$COUNTYFP),
                                 "99999")
```

<br>

## Merging of shapefiles and data

Merge the shape file with the FMR and IL data.

```{r message=FALSE, warning=FALSE}
md_county_fmr_merged = tigris::geo_join(md_county_shapefile,
                                        md_counties_2021_fmr,
                                        by_sp = "fullFips",
                                        by_df = "fips_code",
                                        how = "inner")
```

<br>

## Leaflet mapping

Create the pallete for coloring the overlayed shape file. The scale used
corresponds to Efficiency data.

```{r message=FALSE}
pal <- colorNumeric(
  palette = "Greens",
  domain = as.numeric(md_county_fmr_merged$Efficiency))
```

<br>

Create the county level, efficiency, one bedroom, two bedroom, three bedroom
four bedroom label to appear when hovering using the FMR overlay.

```{r message=FALSE}
label <- paste0(
  "County: ", md_county_fmr_merged$NAME, "<br/>",
  "Efficiency: ", scales::dollar(as.numeric(md_county_fmr_merged$Efficiency)), "<br/>",
  "One Bedroom: ", scales::dollar(as.numeric(md_county_fmr_merged$`One-Bedroom`)), "<br/>",
  "Two Bedroom: ", scales::dollar(as.numeric(md_county_fmr_merged$`Two-Bedroom`)), "<br/>",
  "Three Bedroom: ", scales::dollar(as.numeric(md_county_fmr_merged$`Three-Bedroom`)), "<br/>",
  "Four Bedroom: ", scales::dollar(as.numeric(md_county_fmr_merged$`Four-Bedroom`))) %>% 
  lapply(htmltools::HTML)
```

<br>

Create the ... label to appear when hovering using the IL overlay

``` {r message=FALSE}


```

<br>

Using the leaflet package, plot the county level map of Fair Markets Rents and 
Income Limits in the state of Maryland.

```{r message=FALSE, warning=FALSE, out.width = '100%', fig.height = 6}
md_county_fmr_merged %>% 
  leaflet %>% 
  # add base map
  addProviderTiles("CartoDB") %>% 
  addPolygons(fillColor = ~pal(as.numeric(Efficiency)),
              weight = 2,
              opacity = 1,
              color = "white",
              dashArray = "3",
              fillOpacity = 0.7,
              highlight = highlightOptions(weight = 2,
                                           color = "#666",
                                           dashArray = "",
                                           fillOpacity = 0.7,
                                           bringToFront = TRUE),
              label = label) %>% 
  # add legend
  addLegend(pal = pal, 
            values = ~as.numeric(Efficiency), 
            opacity = 0.7, 
            title = htmltools::HTML("Fair Markets Rent <br> 
                                    by County"),
            position = "bottomright")
```

<br>

## Observations

Some noticable observations are...
